---
title: "Exploring prevalence data and the environmental data"
author: "Himal Shrestha"
date: "03/04/2022"
output: 
  md_document:
    variant: markdown_github
    toc: yes
    toc_depth: 4
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - LA TROBE UNIVERSITY/oncho_gis")
```

## Introduction
This script can be used for following purpose:

1. Filtering raw prevalence data from the [ESPEN database](https://espen.afro.who.int/tools-resources/download-data).

2. Visualising raw onchocerciasis nodule prevalence data along with the map of Ethiopia.

3. Visualise environmental data

4. Extracting environmental data for each data points

5. Exploring environmental data for correlation - Round I selection of environmental variables

## Load libraries
```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
extrafont::loadfonts(device="win", quiet = T)
```

```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
suppressMessages({library(sf);
  library(dplyr);
  library(ggplot2);
  library(rgdal);
  library(leaflet);
  library(raster);
  library(RColorBrewer);
  library(tmap);
  library(gstat);
  library(tidyr);
  library(rasterVis);
  library(cowplot);
  library(climateStability);
  library(car);
  library(GGally);
  library(data.table);
  library(ggpubr);
  library(ggspatial);
  })
```

## Loading the data

```{r}
# data from ESPEN
eth_prev_site <- read.csv("data/data-ET-oncho-sitelevel.csv")

eth_prev_site_2 <- eth_prev_site %>% dplyr::select(ADMIN1_NAME, ADMIN2_NAME,
                                                   IU_NAME, IU_ID, LocationName,
                                                   Longitude, Latitude, 
                                                   Georeliability, SurveyYear,
                                                   Period, Year_MDA1, Method_0,
                                                   Method_2,
                                                   Examined, Positive, 
                                                   Prevalence, CMFL) %>% rename(LONG = Longitude,
                                              LAT = Latitude,
                                              N = Examined,
                                              CASES = Positive)

eth_prev_mf <- eth_prev_site_2 %>% filter(Method_2 == "Parasitology (Skin biopsy)")

eth_prev_nodule <- eth_prev_site_2 %>% filter(Method_2 == "Nodule palpation")
```
+ Finding values from the same locations
```{r}
#df = dataframe object with longitude and latitude coordinates named as `LONG` and `LAT`
find_duplicate <- function(df){
  i <- duplicated(df[, c("LONG", "LAT")])
  long_dup <- df[i, ]$LONG
  lat_dup <- df[i, ]$LAT
  duplicate_sites <- df %>% filter(LONG %in% long_dup & LAT %in% lat_dup)
  return(duplicate_sites)
}
  
duplicate_mf <- find_duplicate(df = eth_prev_mf)
duplicate_nodule <- find_duplicate(df = eth_prev_nodule)
```

+ Aggregating the observations from same locations

```{r}
eth_prev_nodule_2 <- group_by(eth_prev_nodule, LONG, LAT, Method_2) %>%  summarize(
    total = sum(N),
    positive = sum(CASES),
    PREV = positive/total
  )
eth_prev_nodule_2 <- eth_prev_nodule_2 %>% rename(N = total,
                                                  CASES = positive)
eth_prev_nodule_2 %>% find_duplicate()
```
```{r}
eth_prev_mf_2 <- group_by(eth_prev_mf, LONG, LAT, Method_2) %>%
  summarize(
    total = sum(N),
    positive = sum(CASES),
    PREV = positive/total
  )
eth_prev_mf_2 <- eth_prev_mf_2 %>% rename(N = total,
                                              CASES = positive)
```

+ Combining nodule data with unique coordinates: `eth_prev_nodule_2`, and skin mf data `eth_prev_mf_2` with unique coordinates

```{r}
col_required <- names(eth_prev_mf_2)
eth_prev_nodule_2 <- eth_prev_nodule_2[, dput(as.character(names(eth_prev_mf_2)))]
eth_prevalence <- bind_rows(eth_prev_mf_2, eth_prev_nodule_2)
eth_prevalence <- eth_prevalence %>% dplyr::select(LONG, LAT, PREV, Method_2)

eth_prev_same_site <- find_duplicate(df = eth_prevalence)
eth_prev_ss_wide <- eth_prev_same_site %>% pivot_wider(names_from = Method_2,
                                                       values_from = PREV)
```
+ _Correlation analysis__: REMO and skin mf data from the same geographic coordinates were selected for the correlation analysis.

```{r}
cor.test(eth_prev_ss_wide$`Parasitology (Skin biopsy)`, eth_prev_ss_wide$`Nodule palpation`, method = "pearson")

eth_prev_ss_wide$`Parasitology (Skin biopsy)` <- eth_prev_ss_wide$`Parasitology (Skin biopsy)` * 100
eth_prev_ss_wide$`Nodule palpation` <- eth_prev_ss_wide$`Nodule palpation` * 100

cor_plot <- ggscatter(eth_prev_ss_wide, y = "Parasitology (Skin biopsy)", 
          x = "Nodule palpation", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          ylab = "Microfilarial prevalence (%)", xlab = "Nodule prevalence (%)",
          ggtheme = theme_classic(base_family = "Arial"))
cor_plot
```

+ There are two observations which has very high REMO prevalence (>90%) but near to zero mf prevalence. This could be the result of fluctuation in vector dynamics, or human error. These two data were deemed as outliers and not included for the analysis.
+ Removing outliers
```{r}
eth_prev_ss_wide <- eth_prev_ss_wide %>% rename(REMO = "Nodule palpation") 
filter(eth_prev_ss_wide, REMO > 90) -> outlier_obs
outlier_site <- eth_prev_site[eth_prev_site$Longitude %in% outlier_obs$LONG, ]
remove_outlier <- setdiff(eth_prev_ss_wide, outlier_obs)
eth_prev_nodule_3 <- eth_prev_nodule_2[!eth_prev_nodule_2$LONG %in% outlier_obs$LONG, ] %>% mutate(Method_2 = NULL)
```
+ `eth_prev_nodule_3` is the dataset used for the prevalence mapping.

```{r}
# prevalence data
data <- eth_prev_nodule_3
data$prevalence <- data$PREV *100

# Ethiopia border level 0 - country level
m_0 <- getData(name = "GADM", country = "ETH", level = 0)
m_0_sf <- m_0 %>% st_as_sf()

# Ethiopia border level 1 - region level
m_1 <- getData(name = "GADM", country = "ETH", level = 1)
m_1_sf <- m_1 %>% st_as_sf()
```

## Visualize prevalence data on a map
```{r}
mypalette <- colorRampPalette(rev(brewer.pal(11, "RdYlBu")))

# raw prevlaence data on a map
prev <- ggplot() +
  geom_sf(data = m_1_sf, col = "black", lwd = 0.8, lty = 1, alpha = 0) +
  geom_sf(data = m_0_sf, col = "black", lwd = 1, lty = 1, alpha = 0) +
  coord_sf(datum = sf::st_crs(4326))+
  geom_point(data = data, aes(LONG, LAT, fill = prevalence), color = "black", size = 2, stroke = .5, shape = 21, alpha = .7)+ # plotting points with stroke
  theme_void(base_family = "Arial") +
  scale_fill_gradientn(colours = mypalette(100)) +
  xlab("Logitude") + ylab("Latitude") + labs(fill = "Nodule prevalence (%)") + theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'),
                                                                                     #change legend key size
        legend.key.height = unit(.25, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm')) +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5)) +  
  annotation_scale(location = "br", width_hint = 0.5) + #change legend key size 
  annotation_north_arrow(location = "tl", which_north = "true",
                       pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
                       style = north_arrow_fancy_orienteering)
prev

# prevalence histogram
hist(data$prevalence, main = "",
     xlab="Nodule prevalence (%)", ylab = ("Frequency"),
     border="black", 
     col="darkgrey",
     las=1, 
     breaks=30)
```


## Visualise covariate rasters

```{r}
load("data/selected_covs_files_1km.RData")
varlist <- c("slope", "precp_seasonality")
pred_data <- raster::aggregate(selected_covariates[[varlist]], fact = 5, fun = mean, na.rm = TRUE)
```

+ Thanks to the solution from [stack overflow](https://stackoverflow.com/questions/33227182/how-to-set-use-ggplot2-to-map-a-raster)

### Slope
```{r, warning=FALSE}
colr <- colorRampPalette((brewer.pal(9, 'YlOrRd')))
r <- pred_data$slope
my.at <- seq(0, 40, 5)
slope <- levelplot(r, 
          margin=FALSE, xlab = NULL,
          ylab = NULL,
          main = "Slope",
          colorkey=list(
            space='bottom',                   
            labels=list(at=my.at, font=1, height = 0.1), height = 1, width = 0.8,
            axis.line=list(col='black')       
          ),    
          par.settings=list(
            axis.line=list(col='transparent') 
          ),
          scales=list(draw=FALSE),            
          col.regions=colr,                   
          at=seq(0, 40, len=101)) +           
  latticeExtra::layer(sp.polygons(m_0, lwd=1))   
```

### Precipitation seasonality
```{r}
colr <- colorRampPalette((brewer.pal(9, 'PiYG')))
r <- pred_data$precp_seasonality
my.at <- seq(0, 155, 25)
prec_seas <- levelplot(r, 
          margin=FALSE, xlab = NULL,
          ylab = NULL,
          main = "Precipitation seasonality",
          colorkey=list(
            space='bottom',                   
            labels=list(at=my.at, font=1, height = 0.1), height = 1, width = 0.8,
            axis.line=list(col='black')       
          ),    
          par.settings=list(
            axis.line=list(col='transparent') 
          ),
          scales=list(draw=FALSE),            
          col.regions=colr,                   
          at=seq(0, 155, len=101)) +           
  latticeExtra::layer(sp.polygons(m_0, lwd=1))  
```

```{r}
covariate_stack <- plot_grid(prec_seas, slope, labels = "AUTO", nrow = 1)
covariate_stack
```

## Extract environmental values for sample locations

```{r}
covariates <- stack("data/covariates_1_km.grd")
```


```{r}
# Extract the data from the raster--------------------------------------------------------
covs <- raster::extract(covariates, data[c("LONG","LAT")], na.rm = TRUE, df = TRUE)

data_c_covs <- as.data.frame(cbind(data, covs))
# data_c_covs <- na.omit(data_c_covs)
data_c_covs %>% head()
```

## Initial selection of environmental variables

+ Grouping data into related variables
  1. Temperature
  2. Precipitation
  3. Vegetation indices
  4. Hydrological data
  5. Socio-demographic data

```{r}
data_c_covs[, c("N", "CASES", "LONG", "LAT", "ID")] <- NULL
temp_covariates <- c(names(data_c_covs)[2:12], "prevalence")
temp_data <- data_c_covs[, temp_covariates]

precip_covariates <- c(names(data_c_covs)[13:20], "prevalence")
precip_data <- data_c_covs[, precip_covariates]

elev_data <- data_c_covs[, c("alt", "slope", "prevalence")]
vegind_data <- data_c_covs[, c("EVI_2003_11", "NDVI_2003_11", "prevalence")]
hydro_data <- data_c_covs[, c("drain_direction", "Flow_accumulation", "dist_river_DIVA", "Soil_moisture", "prevalence")]
sociodem_data <- data_c_covs[, c("Population_density",
                          "night_lights", "mean_housing_2000_15", "prevalence")]
```

+ Custom functions
```{r}
to_long <- function(data){
  data2 <- data %>% pivot_longer(-prevalence, names_to = "covariates", values_to = "values")
  return(data2)
}
```

### Temperature variables
```{r, warning=FALSE}
temp_data_long <- to_long(data = temp_data)

temp_data$temp_annual_range <- NULL
model1 <- lm(prevalence ~., data = temp_data)  ## `temp_annual_range` is showing problem in the model due to multi-collinearity
vif(model1)
alias(model1)  # use this to check which which covariates are perfectly correlated
```

```{r}
ggpairs(temp_data[, 1:ncol(temp_data)-1])
```
+ Based on the pairs plot except `annual_diurnal_range`, `temp_seasonality`, `isothermality`, all others are correlated with each other. We need to choose one covariate out of those correlated ones. `meantemp_wettestquart` selected because of low VIF.

```{r}
temp_covs_2 <- c("annual_diurnal_range", "temp_seasonality", "isothermality", 
                 "meantemp_wettestquart")
temp_data_covs2 <- data_c_covs[, c(temp_covs_2, "prevalence")]
summary(temp_model2 <- lm(prevalence~., data = temp_data_covs2))
vif(temp_model2)
```

### Precipitation

```{r}
precip_data_long <- to_long(data = precip_data)
precip_model1 <- lm(prevalence ~., data = precip_data);  # summary(precip_model1)
vif(precip_model1)
p <- ggpairs(precip_data[, 1:ncol(precip_data)-1])
p
```
+ Least correlated variables are:
      + Precipitation warmest quarter
      + Precipitation coldest quarter
      + Annual Precipitation, Precipitation Wettest Quarter, Precipitation Wettest Month - Precipitation Wettest Quarter selected based on VIF
      + Precipitation Driest Month, Precipitation Driest Quarter and Precipitation Seasonality - Precipitation Seasonality selected
```{r}
precip_covs_2 <- c("precp_wettest_quart", "precp_seasonality", "precp_warmest_quart", 
                 "precp_coldest_quarter")
precip_data_covs2 <- data_c_covs[, c(precip_covs_2, "prevalence")]
summary(precip_model2 <- lm(prevalence~., data = precip_data_covs2))
vif(precip_model2)
```

### Elevation data
```{r}
elev_data_long <- to_long(data = elev_data)
elev_model1 <- lm(prevalence ~., data = elev_data);  # summary(elev_model1)
vif(elev_model1)
p <- ggpairs(elev_data[, 1:ncol(elev_data)-1])
p
```
+ `alt` and `slope` selected

### Vegetation indices
```{r}
vegind_data_long <- to_long(data = vegind_data)

vegind_model1 <- lm(prevalence ~., data = vegind_data);  # summary(vegind_model1)
vif(vegind_model1)
p <- ggpairs(vegind_data[, 1:ncol(vegind_data)-1])
p
```
+ `EVI` and `NDVI` seems to have almost perfect correlation. `NDVI` selected.

### Hydrological data
```{r}
hydro_data_long <- to_long(data = hydro_data)
hydro_model1 <- lm(prevalence ~., data = hydro_data);  # summary(hydro_model1)
vif(hydro_model1)
p <- ggpairs(hydro_data[, 1:ncol(hydro_data)-1])
p
```
+ `drain direction` not selected while all other selected.

### Sociodemographic data
```{r, warning=FALSE}
sociodem_data_long <- to_long(data = sociodem_data)
sociodem_model1 <- lm(prevalence ~., data = sociodem_data);  # summary(sociodem_model1)
vif(sociodem_model1)
p <- ggpairs(sociodem_data[, 1:ncol(sociodem_data)-1])
p
```

+ All the covariates selected.

## Selected covariates
+ Altogether 17 covariates were selected
```{r}
selected_covs <- c("alt", "annual_diurnal_range", "isothermality", "temp_seasonality", "meantemp_wettestquart",
                   "precp_wettest_quart", "precp_seasonality", "precp_warmest_quart",
                   "precp_coldest_quarter", "NDVI_2003_11",
                   "dist_river_DIVA", "Flow_accumulation",
                   "slope", "Soil_moisture", "night_lights", "mean_housing_2000_15")

selected_covs_data <- data_c_covs[, c(selected_covs, "prevalence")]

selected_covs_data_long <- to_long(data = selected_covs_data)
selected_covs_model1 <- lm(prevalence ~., data = selected_covs_data);  # summary(selected_covs_model1)
vif(selected_covs_model1)
p_corr2 <- ggcorr(selected_covs_data[, 1:ncol(selected_covs_data)-1], palette = "RdBu", label = TRUE)
p_corr2
```
+ altitude and mean temperature wettest quarter are inversely related. Altitude selected as a variable of interest.
+ Isothermality and temperature seasonality are correlated, isothermality selected based on VIF

```{r}
selected_covs <- c("alt", "annual_diurnal_range", "isothermality",
                   "precp_wettest_quart", "precp_seasonality", "precp_warmest_quart",
                   "precp_coldest_quarter", "NDVI_2003_11",
                   "dist_river_DIVA", "Flow_accumulation",
                   "slope", "Soil_moisture", "Population_density", "night_lights", "mean_housing_2000_15")
selected_covs_data <- data_c_covs[, c(selected_covs, "prevalence")]
selected_covs_model1 <- lm(prevalence ~., data = selected_covs_data); 
vif(selected_covs_model1)
```

```{r}
selected_covs_data_2 <- selected_covs_data %>% as.data.table()
col_names <- c("ELV","ADR", "IST", "PWTQ", "PST","PWMQ", "PCQ","NDVI","DTR","FAC","SLP","SLM","PDT", "NLT", "PHI", "PREV")
selected_covs_data_2 <- selected_covs_data_2 %>% setnames(col_names) %>% data.frame()
p_corr2 <- ggcorr(selected_covs_data_2[, 1:ncol(selected_covs_data_2)-1], palette = "RdBu", label = TRUE, name = "Pearson correlation", legend.position = "left", legend.size = "12")
p_corr2 <- p_corr2 + theme(legend.title = element_text(face = "bold"), legend.justification = c(1, 0),
        legend.position = c(0.3, 0.5)) + guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5))
p_corr2
```


