---
title: "Exploring prevalence data and the environmental data"
author: "Himal Shrestha"
date: "22/12/2021"
output: 
  md_document:
    variant: markdown_github
    toc: yes
    toc_depth: 4
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - LA TROBE UNIVERSITY/oncho_git/oncho_gis")
```

## Introduction
This script can be used for following purpose:

1. Visualising raw prevalence data along with the map of Ethiopia

2. Visualise environmental data

3. Extract environmental data for each data points

4. Explore environmental data for correlation - Round I selection of environmental variables

## Load libraries
```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
extrafont::loadfonts(device="win")
```

```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
suppressMessages({library(sf);
  library(dplyr);
  library(ggplot2);
  library(rgdal);
  library(leaflet);
  library(raster);
  library(RColorBrewer);
  library(tmap);
  library(gstat);
  library(tidyr);
  library(rasterVis);
  library(cowplot);
  library(climateStability);
  library(car);
  library(GGally);
  library(data.table);
  })
```

## Loading the data
```{r}
# prevalence data
data <- read.csv("data/eth_espen_nodule_prev.csv")
data$prevalence <- data$CASES/data$N *100

# Ethiopia border level 0 - country level
m_0 <- getData(name = "GADM", country = "ETH", level = 0)
m_0_sf <- m_0 %>% st_as_sf()

# Ethiopia border level 1 - region level
m_1 <- getData(name = "GADM", country = "ETH", level = 1)
m_1_sf <- m_1 %>% st_as_sf()
```

## Visualize prevalence data on a map
```{r}
mypalette <- colorRampPalette(rev(brewer.pal(11, "RdYlBu")))

# raw prevlaence data on a map
prev <- ggplot(m_1_sf) +
  geom_sf()+ coord_sf(datum = sf::st_crs(4326))+
  geom_point(data = data, aes(LONG, LAT, fill = prevalence), color = "black", size = 1.25, stroke = .5, shape = 21, alpha = .7)+ # plotting points with stroke
  theme_bw(base_family = "Verdana") +
  scale_fill_gradientn(colours = mypalette(100)) +
  xlab("Logitude") + ylab("Latitude") + labs(fill = "Nodule prevalence (%)") + theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'),
                                                                                     #change legend key size
        legend.key.height = unit(.25, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'))

prev

# prevalence histogram
prev_hist <- qplot(data$prevalence) + xlab("Nodule prevalence (%)") + ylab("Frequency") + theme_bw(base_family = "Verdana", base_size = 16)

prev_hist
```


## Visualise covariate rasters

```{r}
load("data/selected_covs_files_1km.RData")
varlist <- c("slope", "precp_seasonality")
pred_data <- raster::aggregate(selected_covariates[[varlist]], fact = 5, fun = mean, na.rm = TRUE)
```

+ Thanks to the solution from [stack overflow](https://stackoverflow.com/questions/33227182/how-to-set-use-ggplot2-to-map-a-raster)

### Slope
```{r, warning=FALSE}
colr <- colorRampPalette((brewer.pal(9, 'YlOrRd')))
r <- pred_data$slope
my.at <- seq(0, 40, 5)
slope <- levelplot(r, 
          margin=FALSE, xlab = NULL,
          ylab = NULL,
          main = "Slope",
          colorkey=list(
            space='bottom',                   
            labels=list(at=my.at, font=1, height = 0.1), height = 1, width = 0.8,
            axis.line=list(col='black')       
          ),    
          par.settings=list(
            axis.line=list(col='transparent') 
          ),
          scales=list(draw=FALSE),            
          col.regions=colr,                   
          at=seq(0, 40, len=101)) +           
  latticeExtra::layer(sp.polygons(m_0, lwd=1))   
```

### Precipitation seasonality
```{r}
colr <- colorRampPalette((brewer.pal(9, 'PiYG')))
r <- pred_data$precp_seasonality
my.at <- seq(0, 155, 25)
prec_seas <- levelplot(r, 
          margin=FALSE, xlab = NULL,
          ylab = NULL,
          main = "Precipitation seasonality",
          colorkey=list(
            space='bottom',                   
            labels=list(at=my.at, font=1, height = 0.1), height = 1, width = 0.8,
            axis.line=list(col='black')       
          ),    
          par.settings=list(
            axis.line=list(col='transparent') 
          ),
          scales=list(draw=FALSE),            
          col.regions=colr,                   
          at=seq(0, 155, len=101)) +           
  latticeExtra::layer(sp.polygons(m_0, lwd=1))  
```

```{r}
covariate_stack <- plot_grid(prec_seas, slope, labels = "AUTO", nrow = 1)
covariate_stack
```

## Extract environmental values for sample locations

```{r}
covariates <- stack("data/covariates_1_km.grd")
```


```{r}
# Extract the data from the raster--------------------------------------------------------
covs <- raster::extract(covariates, data[c("LONG","LAT")], na.rm = TRUE, df = TRUE)

data_c_covs <- as.data.frame(cbind(data, covs))
# data_c_covs <- na.omit(data_c_covs)
data_c_covs %>% head()
```

## Initial selection of environmental variables

+ Grouping data into related variables
  1. Temperature
  2. Precipitation
  3. Vegetation indices
  4. Hydrological data
  5. Socio-demographic data

```{r}
data_c_covs[, c("N", "CASES", "LONG", "LAT", "ID")] <- NULL
temp_covariates <- c(names(data_c_covs)[2:12], "prevalence")
temp_data <- data_c_covs[, temp_covariates]

precip_covariates <- c(names(data_c_covs)[13:20], "prevalence")
precip_data <- data_c_covs[, precip_covariates]

elev_data <- data_c_covs[, c("alt", "slope", "prevalence")]
vegind_data <- data_c_covs[, c("EVI_2003_11", "NDVI_2003_11", "prevalence")]
hydro_data <- data_c_covs[, c("drain_direction", "Flow_accumulation", "dist_river_DIVA", "Soil_moisture", "prevalence")]
sociodem_data <- data_c_covs[, c("Population_density",
                          "night_lights", "mean_housing_2000_15", "prevalence")]
```

+ Custom functions
```{r}
to_long <- function(data){
  data2 <- data %>% pivot_longer(-prevalence, names_to = "covariates", values_to = "values")
  return(data2)
}
```

### Temperature variables
```{r, warning=FALSE}
temp_data_long <- to_long(data = temp_data)

temp_data$temp_annual_range <- NULL
model1 <- lm(prevalence ~., data = temp_data)  ## `temp_annual_range` is showing problem in the model due to multi-collinearity
vif(model1)
alias(model1)  # use this to check which which covariates are perfectly correlated
```

```{r}
ggpairs(temp_data[, 1:ncol(temp_data)-1])
```
+ Based on the pairs plot except `annual_diurnal_range`, `temp_seasonality`, `isothermality`, all others are correlated with each other. We need to choose one covariate out of those correlated ones. `meantemp_wettestquart` selected because of low VIF.

```{r}
temp_covs_2 <- c("annual_diurnal_range", "temp_seasonality", "isothermality", 
                 "meantemp_wettestquart")
temp_data_covs2 <- data_c_covs[, c(temp_covs_2, "prevalence")]
summary(temp_model2 <- lm(prevalence~., data = temp_data_covs2))
vif(temp_model2)
```

### Precipitation

```{r}
precip_data_long <- to_long(data = precip_data)
precip_model1 <- lm(prevalence ~., data = precip_data);  # summary(precip_model1)
vif(precip_model1)
p <- ggpairs(precip_data[, 1:ncol(precip_data)-1])
p
```
+ Least correlated variables are:
      + Precipitation warmest quarter
      + Precipitation coldest quarter
      + Annual Precipitation, Precipitation Wettest Quarter, Precipitation Wettest Month - Precipitation Wettest Quarter selected based on VIF
      + Precipitation Driest Month, Precipitation Driest Quarter and Precipitation Seasonality - Precipitation Seasonality selected
```{r}
precip_covs_2 <- c("precp_wettest_quart", "precp_seasonality", "precp_warmest_quart", 
                 "precp_coldest_quarter")
precip_data_covs2 <- data_c_covs[, c(precip_covs_2, "prevalence")]
summary(precip_model2 <- lm(prevalence~., data = precip_data_covs2))
vif(precip_model2)
```

### Elevation data
```{r}
elev_data_long <- to_long(data = elev_data)
elev_model1 <- lm(prevalence ~., data = elev_data);  # summary(elev_model1)
vif(elev_model1)
p <- ggpairs(elev_data[, 1:ncol(elev_data)-1])
p
```
+ `alt` and `slope` selected

### Vegetation indices
```{r}
vegind_data_long <- to_long(data = vegind_data)

vegind_model1 <- lm(prevalence ~., data = vegind_data);  # summary(vegind_model1)
vif(vegind_model1)
p <- ggpairs(vegind_data[, 1:ncol(vegind_data)-1])
p
```
+ The range of `EVI` does not seem to be correct. `NDVI` selected.

### Hydrological data
```{r}
hydro_data_long <- to_long(data = hydro_data)
hydro_model1 <- lm(prevalence ~., data = hydro_data);  # summary(hydro_model1)
vif(hydro_model1)
p <- ggpairs(hydro_data[, 1:ncol(hydro_data)-1])
p
```
+ `drain direction` not selected while all other selected.

### Sociodemographic data
```{r}
sociodem_data_long <- to_long(data = sociodem_data)
sociodem_model1 <- lm(prevalence ~., data = sociodem_data);  # summary(sociodem_model1)
vif(sociodem_model1)
p <- ggpairs(sociodem_data[, 1:ncol(sociodem_data)-1])
p
```

+ All the covariates selected.

## Selected covariates
+ Altogether 17 covariates were selected
```{r}
selected_covs <- c("alt", "annual_diurnal_range", "isothermality", "temp_seasonality", "meantemp_wettestquart",
                   "precp_wettest_quart", "precp_seasonality", "precp_warmest_quart",
                   "precp_coldest_quarter", "NDVI_2003_11",
                   "dist_river_DIVA", "Flow_accumulation",
                   "slope", "Soil_moisture", "night_lights", "mean_housing_2000_15")

selected_covs_data <- data_c_covs[, c(selected_covs, "prevalence")]

selected_covs_data_long <- to_long(data = selected_covs_data)
selected_covs_model1 <- lm(prevalence ~., data = selected_covs_data);  # summary(selected_covs_model1)
vif(selected_covs_model1)
p_corr2 <- ggcorr(selected_covs_data[, 1:ncol(selected_covs_data)-1], palette = "RdBu", label = TRUE)
p_corr2
```
+ altitude and mean temperature wettest quarter are inversely related. Altitude selected as a variable of interest.
+ Isothermality and temperature seasonality are correlated, isothermality selected based on VIF

```{r}
selected_covs <- c("alt", "annual_diurnal_range", "isothermality",
                   "precp_wettest_quart", "precp_seasonality", "precp_warmest_quart",
                   "precp_coldest_quarter", "NDVI_2003_11",
                   "dist_river_DIVA", "Flow_accumulation",
                   "slope", "Soil_moisture", "Population_density", "night_lights", "mean_housing_2000_15")
selected_covs_data <- data_c_covs[, c(selected_covs, "prevalence")]
selected_covs_model1 <- lm(prevalence ~., data = selected_covs_data); 
vif(selected_covs_model1)
```

```{r}
selected_covs_data_2 <- selected_covs_data %>% as.data.table()
col_names <- c("ELV","ADR", "IST", "PWTQ", "PST","PWMQ", "PCQ","NDVI","DTR","FAC","SLP","SLM","PDT", "NLT", "PHI", "PREV")
selected_covs_data_2 <- selected_covs_data_2 %>% setnames(col_names) %>% data.frame()
p_corr2 <- ggcorr(selected_covs_data_2[, 1:ncol(selected_covs_data_2)-1], palette = "RdBu", label = TRUE)
p_corr2
```


