---
title: "Model Selection and Fitting"
author: "Himal Shrestha"
date: "22/12/2021"
output: 
  md_document:
    variant: markdown_github
    toc: yes
    toc_depth: 4
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - LA TROBE UNIVERSITY/oncho_git/oncho_gis")
```

## Introduction
This script can be used for following purpose:

1. Testing different distributions for modelling the prevalence data.

2. Testing different triangulation meshes

3. Running univariate model for the different environmental variables

4. K-fold cross validation of the candidate models

## Load libraries
```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
extrafont::loadfonts(device="win")
```

```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
suppressMessages({library(tidyverse);
  library(INLA);
  library(INLAutils);
  library(raster);
  library(sf);
  library(dplyr);
  library(tictoc);
  library(leaflet);
  library(GGally);
  library(devtools);
  library(sp);
  library(cowplot);
  library(ggplot2);
  })
```

+ custom function for stacking the data
```{r}
# data = data from sample locations includes prevalence data as CASES and N, and includes the values for environmental variables in the sample locations
# dp = data for the locations to be predicted. It includes all the environmental data for the locations where the prevalence data is to be predicted. The CASES and N column for dp is blank as these data are the ones to be predicted
# cov_list = vector of selected covariates for the model fitting and prediction
stack_data <- function(data, dp, cov_list){
  # stack for estimation stk.e
  df <- data.frame(Intercept = 1, subset(data, select = cov_list))
  stk.e <- inla.stack(
    tag = "est",
    data = list(y = data$CASES, numtrials = data$N),
    A = list(1, Ae),
    effects = list(df, spatial.field = iset)
  )
  
  # stack for prediction stk.p
  df_p <- data.frame(Intercept = 1, subset(dp, select = cov_list))
  stk.p <- inla.stack(
    tag = "pred",
    data = list(y = dp$CASES, numtrials = dp$N),
    A = list(1, Ap),
    effects = list(df_p, spatial.field = iset))
  
  # stk.full has stk.e and stk.p
  stk.full <- inla.stack(stk.e, stk.p)
  
  return(stk.full)
}
```

## Loading the data

```{r}
load("data/selected_covs_files_1km.RData")
data <- eth_prev2

m <- getData(name = "GADM", country = "ETH", level = 0)
# m <- m %>% st_as_sf()
# pred_data <- selected_covariates
# ra <- pred_data
```

+ Test mesh for model selection
```{r}
coords <-  cbind(data$LONG, data$LAT)
bdry <- inla.sp2segment(m)
bdry$loc <- inla.mesh.map(bdry$loc)

### Mesh 1
mesh1 <- inla.mesh.2d(
  loc = coords, boundary = bdry, 
  max.edge = c(0.5, 5),
  cutoff = 0.3
)
mesh1$n
mesh1_plot <- autoplot(mesh1) + theme_void() + theme(legend.position = "none")
mesh1_plot
```
### SPDE and projection matrix

```{r}
A <- inla.spde.make.A(mesh = mesh1, loc = as.matrix(coords)); dim(A)
spde <- inla.spde2.matern(mesh1, alpha=2)
# spde <- inla.spde2.pcmatern(mesh = mesh1, alpha = 2, constr = TRUE,
#                             prior.range = c(15, 0.01), #P(range < 15000) = 0.01
#                             prior.sigma = c(3, 0.01))  # P(sigma > 3) = 0.01

iset <- inla.spde.make.index(name = "spatial.field", spde$n.spde)
length(iset)
```
### INLA stack
```{r}
df <- data.frame(Intercept = 1, subset(data, select = selected_covs))

stk <- inla.stack(
    tag = "est",
    data = list(y = data$CASES, numtrials = data$N),
    A = list(1, A),
    effects = list(df, spatial.field = iset)
  )
```

### Model fitting
```{r}
formula01 <-y ~ -1 + Intercept  # without spatial effect
formula0 <-y ~ -1 + Intercept + f(spatial.field, model=spde)  # with spatial effect
# + Soil_moisture + Flow_accumulation + near_river_DIVA + night_lights + mean_housing_2000_15 + Population_density + annual_diurnal_range + annual_precp + f(spatial.field, model=spde) 
```

#### Binomial
- Without spatial effect
```{r}
tic()
inla.setOption(num.threads = 12)
model01.binom <- inla(formula01,
            family = "binomial", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)
            )
  )
toc()
```

- With spatial effect
```{r}
tic()
inla.setOption(num.threads = 12)
model.binom <- inla(formula0,
            family = "binomial", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)
            )
  )
toc()
```

#### Zero inflated binomial type I
```{r}
tic()
inla.setOption(num.threads = 12)
model.zib0 <- inla(formula0,
            family = "zeroinflated.binomial.0", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)
            )
  )
toc()

```
#### Zero inflated binomial type II
```{r}
set.seed(12345)
tic()
inla.setOption(num.threads = 12)
model.zib1 <- inla(formula0,
            family = "zeroinflated.binomial.1", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)
            )
  )
toc()
```

```{r}
model_stats <- tibble(Models = c("Binomial without spatial","Binomial", "Zero Inflated Binomial Type 0",
                  "Zero Inflated Binomial Type 1"),
       DIC = c(model01.binom$dic$dic, model.binom$dic$dic, model.zib0$dic$dic, model.zib1$dic$dic),
       WAIC = c(model01.binom$waic$waic, model.binom$waic$waic, model.zib0$waic$waic, model.zib1$waic$waic)) 
model_stats

model_stats_long <- model_stats %>% mutate(Models = as_factor(Models)) %>%
  pivot_longer(cols = c("DIC", "WAIC"),
               names_to = "validation_stats",
               values_to = "values") %>% mutate(validation_stats = as_factor(validation_stats))


p <- ggplot(model_stats_long, aes(x=Models, y=values, group=validation_stats)) +
  geom_line(aes(linetype=validation_stats, color = validation_stats), size = 1.3)+
  geom_point(aes(fill = validation_stats), show.legend = TRUE, size = 2, shape = 21, stroke = 1)+
  scale_linetype_manual(values=c("solid", "solid")) +
  scale_color_manual(values = c("#E69F00", "skyblue")) +
  scale_fill_manual(values = c("#E69F00", "skyblue")) +
theme_bw(base_family = "Verdana", base_size = 12) +
  labs(y = "Validation statistics", color = "", linetype = "", fill = "") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) #+ facet_wrap(~validation_stats, scales = "free") +
p
```

+ This shows that prevalence data can be best modeled with zero inflated type I distribution.

## Testing different triangulation meshes

```{r}
### mesh 2
mesh2 <- inla.mesh.2d(
  loc = coords, boundary = bdry, 
  max.edge = c(0.5, 5),
  cutoff = 0.1
)
mesh2$n
mesh2_plot <- autoplot(mesh2) + theme_void() + theme(legend.position = "none")

## mesh 3 - Paula's mesh
mesh3 <- inla.mesh.2d(
  loc = coords, boundary = bdry, max.edge = c(0.1, 5),
  cutoff = 0.01
)
mesh3$n
mesh3_plot <- autoplot(mesh3) + theme_void() + theme(legend.position = "none")


### mesh 4 - changing the triangles outside bounday
mesh4 <- inla.mesh.2d(
  loc = coords, boundary = bdry, max.edge = c(0.5, 1),
  cutoff = 0.1
)
mesh4$n
mesh4_plot <- autoplot(mesh4) + theme_void() + theme(legend.position = "none")


### mesh 5
mesh5 <- inla.mesh.2d(
  loc = coords, boundary = bdry, max.edge = c(0.5, 5),
  cutoff = 0.03
)
mesh5$n
mesh5_plot <- autoplot(mesh5) + theme_void() + theme(legend.position = "none")


### mesh 6 - changing the triangles inside boundary
mesh6 <- inla.mesh.2d(
  loc = coords, boundary = bdry, max.edge = c(0.3, 5),
  cutoff = 0.1
)
mesh6$n
mesh6_plot <- autoplot(mesh6) + theme_void() + theme(legend.position = "none")

meshes <- plot_grid(mesh1_plot, mesh2_plot, mesh3_plot, mesh4_plot, mesh5_plot, mesh6_plot, labels = "AUTO")

meshes
```

#### Running model on different meshes
```{r, eval = FALSE}
df_1 <- data.frame(Meshes = 1:6,
                   DIC = NA,
                   WAIC = NA,
                   time_taken = NA,
                   number_p = NA)  # create dummy data frame to store model fit scores

meshes <- list(mesh1, mesh2, mesh3, mesh4, mesh5, mesh6)

pb = txtProgressBar(min = 0, max = length(meshes), initial = 0) 

for (i in 1:length(meshes)) {
  A <- inla.spde.make.A(mesh = meshes[[i]], loc = as.matrix(coords)); dim(A)
  spde <- inla.spde2.matern(meshes[[i]], alpha=2)
  iset <- inla.spde.make.index(name = "spatial.field", spde$n.spde)
  
  stk <- inla.stack(
    tag = "est",
    data = list(y = data$CASES, numtrials = data$N),
    A = list(1, A),
    effects = list(df, spatial.field = iset)
  )
  
  inla.setOption(num.threads = 12)
  old <- Sys.time()
  p.res <- inla(formula0,
            family = "zeroinflated.binomial.1", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)
            )
  )
  df_1[i, 4] <- Sys.time() - old
  df_1[i, 2] <- p.res$dic$dic
  df_1[i, 3] <- p.res$waic$waic
  df_1[i, 5] <- meshes[[i]]$n
  
  setTxtProgressBar(pb,i)
  cat("\nIteration = ", i, "\n")
  print(Sys.time() - old)
}

df_1
```
## Selection of environmental variables
```{r}
mesh <- mesh5
A <- inla.spde.make.A(mesh = mesh, loc = as.matrix(coords)); dim(A)
spde <- inla.spde2.matern(mesh, alpha=2)
iset <- inla.spde.make.index(name = "spatial.field", spde$n.spde)

df <- data.frame(Intercept = 1, subset(data, select = selected_covs))
stk <- inla.stack(
    tag = "est",
    data = list(y = data$CASES, numtrials = data$N),
    A = list(1, A),
    effects = list(df, spatial.field = iset)
  )
```

```{r}
covs_all <- names(data)[6:20]
```

### Univariate model
```{r, eval = FALSE}
covs <- covs_all
pb = txtProgressBar(min = 0, max = length(covs[1:15]), initial = 0) 

old <- Sys.time()
for(i in 1:15){
  set.seed(12345)
  f1 <- as.formula(paste0("y ~ -1 + Intercept + f(spatial.field, model=spde) + ", 
                          paste0(covs[i], collapse = " + ")))
  inla.setOption(num.threads = 12)
  model1<-inla(f1,
            family = "zeroinflated.binomial.1", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)))
  model_selection <- if(i==1){rbind(c(model = paste(covs[i]),
                                      waic =model1$waic$waic,
                                      dic = model1$dic$dic))}
  else{rbind(model_selection,c(model = paste(covs[i],collapse = " + "),
                               waic = model1$waic$waic, dic = model1$dic$dic
                               ))}
  setTxtProgressBar(pb,i)
  cat("\nIteration = ", i, "\n")
  print(Sys.time() - old)
}

model_selection <- data.frame(model_selection)
model_selection
```

### Other models
```{r, warning=FALSE}
# variables from each group that yielded the least DIC
covs_1 <- c("slope", "isothermality", "precp_seasonality", "NDVI_2003_11", "Population_density", "dist_river_DIVA")
# testing other combinations
covs_2 <- c("slope", "isothermality", "precp_seasonality", "NDVI_2003_11", "Population_density", "dist_river_DIVA", "Flow_accumulation", "Soil_moisture")

covs_comb <- list(covs_1, covs_2)

pb = txtProgressBar(min = 0, max = length(covs_comb), initial = 0) 

old <- Sys.time()
for(i in 1:length(covs_comb)){
  set.seed(12345)
  f1 <- as.formula(paste0("y ~ -1 + Intercept + f(spatial.field, model=spde) + ", 
                          paste0(covs_comb[[i]], collapse = " + ")))
  inla.setOption(num.threads = 12)
  model1<-inla(f1,
            family = "zeroinflated.binomial.1", Ntrials = numtrials,
            data = inla.stack.data(stk, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(stk)))
  model_selection <- if(i==1){rbind(c(model = paste(covs_comb[[i]],collapse = " + "),
                                      waic =model1$waic$waic,
                                      dic = model1$dic$dic))}
  else{rbind(model_selection,c(model = paste(covs_comb[[i]],collapse = " + "),
                               waic = model1$waic$waic, dic = model1$dic$dic
                               ))}
  setTxtProgressBar(pb,i)
  cat("\nIteration = ", i, "\n")
  print(Sys.time() - old)
}

model_selection <- data.frame(model_selection)
model_selection
```

## K-fold cross validation between the candidate models

### Candidate models
```{r}
varlist <- c("slope", "isothermality", "precp_seasonality", "NDVI_2003_11",
             "dist_river_DIVA", "Flow_accumulation",
             "Soil_moisture", "Population_density")

formula0 <- as.formula("y ~ 0 + Intercept + f(spatial.field, model = spde)")
formula1 <- as.formula(paste("y ~ 0 + Intercept +", paste(varlist[!varlist %in% c("Flow_accumulation", "Soil_moisture")], collapse =  "+"), "+ f(spatial.field, model = spde)"))
formula2 <- as.formula(paste("y ~ 0 + Intercept +", paste(varlist, collapse =  "+"), "+ f(spatial.field, model = spde)"))

formula <- list(formula0, formula1, formula2)

data$observed <- data$CASES
data$PREV <- data$CASES/data$N
```

### Model validation
+ Output RMSE and correlation ceofficient
+ Output validation data with observed and predicted values
```{r, eval = FALSE}
library(dismo)
fold <- 10 # specify fold
# model <- 2 # specify model
variables = 2
pb = txtProgressBar(min = 0, max = fold, initial = 0) 
output <- matrix(ncol=variables+2)
colnames(output) <- c("Model", "Fold", "RMSE", "Correlation-coefficient")
output <- data.frame(output)

set.seed(12345)
kf <- kfold(nrow(data), k = fold)

old <- Sys.time()
for (j in 1:length(formula)) {
  for(i in 1:fold) {
    test <- data[kf == i, ]
    train <- data[kf != i, ]
    test$CASES <- NA  #make the y values for test NA
    test_coords <- coords[kf == i,]
    train_coords <- coords[kf != i,]
    Ae <- inla.spde.make.A(mesh=mesh5,loc=as.matrix(train_coords));dim(Ae)
    Ap <- inla.spde.make.A(mesh = mesh5, loc = test_coords);dim(Ap)
    stk.full <- stack_data(data = train, dp = test, cov_list = selected_covs)
    inla.setOption(num.threads = 8)
    p.res <- inla(formula[[j]],
              family = "zeroinflated.binomial.1", Ntrials = numtrials,
              data = inla.stack.data(stk.full, spde = spde),
              control.family = list(link = "logit"),
              control.compute = list(dic = TRUE, waic = TRUE,
                                     cpo = TRUE, config = TRUE,                                     openmp.strategy="huge"),
              control.predictor = list(
                compute = TRUE, link = 1,
                A = inla.stack.A(stk.full)
              )
    )
    index.pred <- inla.stack.index(stk.full, "pred")$data
    obs_prev <- test$PREV #this is the number pos/number examined
    tmp.mean <- p.res$summary.fitted.values[index.pred, "mean"]
    tmp.sd = p.res$summary.fitted.values[index.pred,"sd"]
    validation = list()
    validation$res = obs_prev - tmp.mean 
    validation$rmse = sqrt(mean(validation$res^2, na.rm=TRUE)) 
    validation$cor = cor(obs_prev, tmp.mean, 
                         use="pairwise.complete.obs",
                         method="spearman")
    output <- bind_rows(output, (data.frame(Model = j, Fold = i, RMSE = validation$rmse, Correlation.coefficient = validation$cor))) 
    
    if(i == 1){
        object_cor <- data.frame(model = j, fold = i, tmp.mean, obs_prev)
      }
      else{
        object_cor <- bind_rows(object_cor, data.frame(model = j, fold = i, tmp.mean, obs_prev))
      }
      setTxtProgressBar(pb,i)
      cat("\nIteration = ", i, "\n")
      print(Sys.time() - old)
  }
  if (j == 1) {
    object_cor_combined <- object_cor
  }
  else{
    object_cor_combined <- bind_rows(object_cor_combined, object_cor)
  }
  cat(paste(c("Model ", j, " completed"), collapse = ""))
}
```

+ This is the script to run k fold cross-validation for different `k` and different `formula`/models
+ The output that we are interested in are:
1. `output` - dataframe with RMSE and R-squared values for different models and folds
2. `object_cor_combined` - dataframe with observed and predicted prevalence values in each folds of cross-validation for different models

```{r}
load("data/211231_output_cross_val.RData")

output %>% head()

object_cor_combined %>% head()

output <- output %>% mutate(Model = as_factor(Model))
output$model_names <- rep(c("Model 0", "Model 1", "Model 2"), each = 10)
output <- output %>% mutate(model_names = as_factor(model_names))
```

#### Plot RMSE and correlation coefficient
```{r}
library(forcats)

cor_coff <- output %>% mutate(model_names = fct_reorder(model_names, Correlation.coefficient)) %>% 
  ggplot( aes(x=model_names, y=Correlation.coefficient, fill = model_names, alpha = 0.7)) +
    geom_boxplot(outlier.shape = NA, width = 0.6) +
    geom_point(shape = 21, fill = "white", color = "black", size = 2, alpha = 0.5) + #geom_jitter(color="black", size=2, alpha=0.9) +
    theme_bw(base_family = "Verdana") +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    xlab("Models") + ylab("Correlation coefficient") + ylim(c(0.3,.8))

output$RMSE <- output$RMSE * 100  # multiplied as prevalence is expressed as percentage in the model output

RMSE <- output %>% mutate(model_names = fct_reorder(model_names, RMSE, .desc = T)) %>% 
  ggplot( aes(x=model_names, y=RMSE, fill = model_names, alpha = 0.7)) +
    geom_boxplot(outlier.shape = NA, width = 0.6) +
    geom_point(shape = 21, fill = "white", color = "black", size = 2, alpha = 0.5) + #geom_jitter(color="black", size=2, alpha=0.9) +
    theme_bw(base_family = "Verdana") +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    xlab("Models") + ylab("RMSE") + ylim(c(8,16))

CV_plot <- plot_grid(cor_coff, RMSE, nrow = 1, labels = "AUTO")

CV_plot
```

+ This concludes that the `Model 2` is the best model which has the following formula:

```
y ~ 0 + Intercept + slope + isothermality + precp seasonality + 
    NDVI_2003_11 + dist_river_DIVA + Flow_accumulation + Soil_moisture + 
    Population_density + f(spatial.field, model = spde)
```
