---
title: "Model Selection and Fitting"
author: "Himal Shrestha"
date: "22/12/2021"
output: 
  md_document:
    variant: markdown_github
    toc: yes
    toc_depth: 4
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - LA TROBE UNIVERSITY/oncho_gis")
```

## Introduction
This script can be used for following purpose:

1. Running the best model using INLA

2. Prediction using the model to create predicted prevalence map across Ethiopia

3. Creating uncertainty bivariate map

4. Creating exceedance probability map

5. Aggregated map based on health districts: mean prevalence, range, case numbers.

6. Relationship profile between environmental variables and the predicted prevalence 

7. Running the regular binomial model 

## Load libraries
```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
extrafont::loadfonts(device="win", quiet = TRUE)
```

```{r, message=FALSE, error=FALSE, tidy=TRUE, warning=FALSE}
suppressMessages({library(tidyverse);
  library(INLA);
  library(INLAutils);
  library(raster);
  library(sf);
  library(dplyr);
  library(tictoc);
  library(leaflet);
  library(GGally);
  library(devtools);
  library(sp);
  library(cowplot);
  library(tmap);
  library(RColorBrewer);
  library(patchwork);
  ## for bivariate maps
  library(classInt);
  library(rgdal);
  library(dismo);
  library(XML);
  library(maps);
  library(sp);
  library(ggspatial);
  })
```

+ custom function for stacking the data
```{r}
# data = data from sample locations includes prevalence data as CASES and N, and includes the values for environmental variables in the sample locations
# dp = data for the locations to be predicted. It includes all the environmental data for the locations where the prevalence data is to be predicted. The CASES and N column for dp is blank as these data are the ones to be predicted
# cov_list = vector of selected covariates for the model fitting and prediction
stack_data <- function(data, dp, cov_list){
  # stack for estimation stk.e
  df <- data.frame(Intercept = 1, subset(data, select = cov_list))
  stk.e <- inla.stack(
    tag = "est",
    data = list(y = data$CASES, numtrials = data$N),
    A = list(1, Ae),
    effects = list(df, spatial.field = iset)
  )
  
  # stack for prediction stk.p
  df_p <- data.frame(Intercept = 1, subset(dp, select = cov_list))
  stk.p <- inla.stack(
    tag = "pred",
    data = list(y = NA, numtrials = NA),
    A = list(1, Ap),
    effects = list(df_p, spatial.field = iset))
  
  # stk.full has stk.e and stk.p
  stk.full <- inla.stack(stk.e, stk.p)
  
  return(stk.full)
}
```

## Loading the data

```{r}
load("data/selected_covs_files_1km.RData")
data <- eth_prev2

m <- getData(name = "GADM", country = "ETH", level = 0)
m_0 <- m %>% st_as_sf()
# pred_data <- selected_covariates
# ra <- pred_data

covariates <- selected_covariates
varlist <- c("slope", "isothermality", "precp_seasonality",
             "NDVI_2003_11", "dist_river_DIVA", "Population_density",
             "Flow_accumulation", "Soil_moisture") 
pred_data <- covariates[[varlist]]
pred_data <- aggregate(covariates[[varlist]], fact = 5, fun = mean, na.rm = TRUE) ## aggregating the prediction grid to 5 x 5 km grid

ra <- pred_data
dp <- data.frame(rasterToPoints(ra))
```

## Model fitting
### Mesh construction
```{r}
coords <-  cbind(data$LONG, data$LAT)
bdry <- inla.sp2segment(m)
bdry$loc <- inla.mesh.map(bdry$loc)
mesh1 <- inla.mesh.2d(
  loc = coords, boundary = bdry, max.edge = c(0.5, 5),
  cutoff = 0.03 # best mesh
)
mesh1$n

spde <- inla.spde2.matern(mesh1, alpha=2)
iset <- inla.spde.make.index(name = "spatial.field", spde$n.spde)
length(iset)
```

### Make projection matrix
```{r}
Ae <- inla.spde.make.A(mesh=mesh1,loc=as.matrix(coords));dim(Ae)
coop <- cbind(dp$x, dp$y)
Ap <- inla.spde.make.A(mesh = mesh1, loc = coop);dim(Ap)
```

### Model formula
```{r}
predterms <- as.formula(paste("y ~ 0 + Intercept +", paste(varlist, collapse =  "+"), "+ f(spatial.field, model = spde)"))
datastack <- stack_data(data = data, dp = dp, cov_list = varlist)
```

### Model fitting
```{r, eval = FALSE}
tic()
res <- inla(predterms,
            family = "zeroinflated.binomial.1", Ntrials = numtrials,
            data = inla.stack.data(datastack, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(datastack)
            )
)
toc()
```

### Model results
```{r}
res <- readRDS("data/inla_point_res_5km_mesh5.rds")
summary(res)
reg_coff <- res$summary.fixed %>% arrange(desc(mean))

### Fixed effects from the model
reg_coff %>% round(digits = 5)

### Hyper parameters from the model
res$summary.hyperpar %>% data.frame() %>% round(digits = 5)
```

## Model prediction
### Mapping prediction
```{r}
index <- inla.stack.index(stack = datastack, tag = "pred")$data

prev_mean <- res$summary.fitted.values[index, "mean"]
prev_ll <- res$summary.fitted.values[index, "0.025quant"]
prev_ul <- res$summary.fitted.values[index, "0.975quant"]
prev_sd <-  res$summary.fitted.values[index, "sd"]

summary(res$summary.fitted.values[index,])
```

```{r}
load("data/220507_ZIB1_maps.RData")
### Custom function to rasterize the predicted prevalence
raster_prev <- function(prev_mean){
  r_prev_mean <- rasterize(
    x = coop, y = ra, field = prev_mean,
    fun = mean)
}
```

```{r, eval = FALSE}
r_prev_mean <- raster_prev(prev_mean = prev_mean)
r_prev_ul <- raster_prev(prev_ul)
r_prev_ll <- raster_prev(prev_ll)
r_prev_sd <- raster_prev(prev_sd)
```

### Prevalence map
```{r}
pal <- hcl.colors(100, palette = "Reds", rev =TRUE)
my.palette.heat <- rev(brewer.pal(n = 9, name = "RdYlBu"))


p_mean <- tm_shape(r_prev_mean*100) + 
    tm_raster(title="Mean Prevalence (%)", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 50) +
    tm_shape(m_0)+
    tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
    tm_compass(type = "arrow", position = c("left", "top")) +
    tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
    tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p_mean


p_UL <- tm_shape(r_prev_ul*100) + 
  tm_raster(title="Prevalence (%): Upper limit", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 50) +
  tm_shape(m_0)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  # tm_compass(type = "8star", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")


p_LL <- tm_shape(r_prev_ll*100) + 
  tm_raster(title="Prevalence (%): Lower limit", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 50) +
  tm_shape(m_0)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  # tm_compass(type = "8star", position = c("left", "top")) +
  # tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")

tmap_arrange(p_LL, p_UL, ncol = 1)
```

### Uncertainty map
```{r}
data_sf <- st_as_sf(x = data, coords = c('LONG','LAT'), crs= "+proj=longlat +datum=WGS84")

pal <- hcl.colors(100, palette = "Blues", rev =TRUE)
p <- tm_shape(r_prev_sd*100) + 
  tm_raster(title="Prevalence(SD)", alpha = .8, palette = pal, legend.is.portrait = T, style = "cont") +
  tm_shape(m_0)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_shape(data_sf) + 
  tm_symbols(col = "black", shape = 3, alpha = .5, size = .07, legend.shape.show = T) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p
```

### Bivariate map
+ Thanks to this [blog](https://rfunctions.blogspot.com/2015/03/bivariate-maps-bivariatemap-function.html)
#### Bivariate map functions
```{r}
colmat<-function(nquantiles=10, upperleft=rgb(0,150,235, maxColorValue=255), upperright=rgb(130,0,80, maxColorValue=255), bottomleft="grey", bottomright=rgb(255,230,15, maxColorValue=255), xlab="x label", ylab="y label"){
  my.data<-seq(0,1,.01)
  my.class<-classIntervals(my.data,n=nquantiles,style="quantile")
  my.pal.1<-findColours(my.class,c(upperleft,bottomleft))
  my.pal.2<-findColours(my.class,c(upperright, bottomright))
  col.matrix<-matrix(nrow = 101, ncol = 101, NA)
  for(i in 1:101){
    my.col<-c(paste(my.pal.1[i]),paste(my.pal.2[i]))
    col.matrix[102-i,]<-findColours(my.class,my.col)}
  plot(c(1,1),pch=19,col=my.pal.1, cex=0.5,xlim=c(0,1),ylim=c(0,1),frame.plot=F, xlab=xlab, ylab=ylab,cex.lab=1.3)
  for(i in 1:101){
    col.temp<-col.matrix[i-1,]
    points(my.data,rep((i-1)/100,101),pch=15,col=col.temp, cex=1)}
  seqs<-seq(0,100,(100/nquantiles))
  seqs[1]<-1
  col.matrix<-col.matrix[c(seqs), c(seqs)]}
```

```{r}
bivariate.map<-function(rasterx, rastery, colormatrix=col.matrix, nquantiles=10){
  quanmean<-getValues(rasterx)
  temp<-data.frame(quanmean, quantile=rep(NA, length(quanmean)))
  brks<-with(temp, quantile(temp,na.rm=TRUE, probs = c(seq(0,1,1/nquantiles))))
  r1<-within(temp, quantile <- cut(quanmean, breaks = brks, labels = 2:length(brks),include.lowest = TRUE))
  quantr<-data.frame(r1[,2]) 
  quanvar<-getValues(rastery)
  temp<-data.frame(quanvar, quantile=rep(NA, length(quanvar)))
  brks<-with(temp, quantile(temp,na.rm=TRUE, probs = c(seq(0,1,1/nquantiles))))
  r2<-within(temp, quantile <- cut(quanvar, breaks = brks, labels = 2:length(brks),include.lowest = TRUE))
  quantr2<-data.frame(r2[,2])
  as.numeric.factor<-function(x) {as.numeric(levels(x))[x]}
  col.matrix2<-colormatrix
  cn<-unique(colormatrix)
  for(i in 1:length(col.matrix2)){
    ifelse(is.na(col.matrix2[i]),col.matrix2[i]<-1,col.matrix2[i]<-which(col.matrix2[i]==cn)[1])}
  cols<-numeric(length(quantr[,1]))
  for(i in 1:length(quantr[,1])){
    a<-as.numeric.factor(quantr[i,1])
    b<-as.numeric.factor(quantr2[i,1])
    cols[i]<-as.numeric(col.matrix2[b,a])}
  r<-rasterx
  r[1:length(r)]<-cols
  return(r)}
```

#### Plot bivariate map

```{r}
# Color matrix
col.matrix <-colmat(nquantiles=4, upperleft="lightgreen", upperright="#F0842B5E", 
                   bottomleft="#6EA2F0", bottomright="#B31212D4", xlab="Prevalence", ylab="Uncertainty" )
```

```{r}
# Bivariate map
bivmap<-bivariate.map(r_prev_mean,r_prev_sd, colormatrix=col.matrix, nquantiles=4)
plot(bivmap,frame.plot=F,axes=F,box=F,add=F,legend=F,col=as.vector(col.matrix))
plot(st_geometry(m_0), add= T, lwd = 4)
```

### Creating exceedance probability map

```{r, eval = FALSE}
excprob_20 <- sapply(res$marginals.fitted.values[index],
                  FUN = function(marg){1-inla.pmarginal(q = 0.20, marginal = marg)})
raster_excprob_20 <- raster_prev(excprob_20)
```

```{r}
pal <- hcl.colors(100, palette = "RdYlBu", rev =TRUE)


excprob_20_map <- tm_shape(raster_excprob_20) + 
  tm_raster(title="P (nodule prevalence > 20%)", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 0.5) +
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
excprob_20_map
```

### Predicting case numbers
```{r}
eth_pop_count2012 <- raster("data/eth_ppp_2012_1km_Aggregated_UNadj.tif")
eth_pop_count2012 %>% values() %>% sum(na.rm = t)  # actual value

# custom function for reprojection
reproj <- function(tbc, c){
  tbc <- projectRaster(tbc, c)
  tbc <- resample(tbc, c)
  tbc <- mask(tbc, c)
}

load("data/220507_case_numbers.RData")
```

```{r, eval = FALSE}
eth_pop_count <- reproj(eth_pop_count2012, covariates[[1]])
names(eth_pop_count) <- "eth_popcount_2012"
```

```{r}
eth_popcount_5km <-  aggregate(eth_pop_count, fact = 5, fun = sum, na.rm = TRUE) # aggregating to 5 km
```

+ Excluding urban areas
```{r}
# Custom function for reporjective categorical rasters
reproj_cat <- function(tbc, c){
  tbc <- projectRaster(tbc, c, method = "ngb")
  tbc <- resample(tbc, c, method = "ngb")
  tbc <- mask(tbc, c)
}
```

```{r}
rur_urb_eth2015 <- raster("data/220401_rurban_eth_2015_1km.tif")
rur_urb_eth2015 %>% values() %>% as.factor() %>% summary()
```

```{r, eval = FALSE}
rururb_eth2015_reproj <- reproj_cat(rur_urb_eth2015, covariates[[1]])
rururb_eth2015_reproj %>% values() %>% as.factor() %>% summary()
```

```{r}
rur_urb_eth2015_reproj_5km <-  aggregate(rururb_eth2015_reproj, fact = 5, fun = modal, na.rm = TRUE)
rur_urb_eth2015_reproj_5km %>% values() %>% as.factor() %>% summary()

# 2: semi-urban; 3: urban
rururb_nourban_5km <- reclassify(rur_urb_eth2015_reproj_5km, c(-Inf, 0, 1, 
                                                               0, 1, 1,
                                                                  1, Inf, NA))
rururb_nourban_5km[] %>% as.factor %>% summary
```

```{r}
eth_popcount2012_5km_nourban <- eth_popcount_5km * rururb_nourban_5km
eth_popcount2012_5km_nourban %>% values() %>% sum(na.rm = T)
# Number of cases
mean_case <- eth_popcount2012_5km_nourban * r_prev_mean
mean_case %>% values() %>% sum(na.rm = T)

LL_case <- eth_popcount2012_5km_nourban * r_prev_ll
LL_case %>% values() %>% sum(na.rm = T)

UL_case <- eth_popcount2012_5km_nourban * r_prev_ul
UL_case %>% values() %>% sum(na.rm = T)
```

```{r}
eth_st2 <- getData(name = "GADM", country = "ETH", level = 3)
eth_sf2 <- eth_st2 %>% st_as_sf()

mean_cases <- exactextractr::exact_extract(mean_case, eth_sf2, fun = "sum") # faster
eth_sf2$mean_cases <- mean_cases/1000

LL_cases <- exactextractr::exact_extract(LL_case, eth_sf2, fun = "sum") # faster
eth_sf2$LL_cases <- LL_cases/1000

UL_cases <- exactextractr::exact_extract(UL_case, eth_sf2, fun = "sum") # faster
eth_sf2$UL_cases <- UL_cases/1000
```

```{r}
mypalette <- colorRampPalette((brewer.pal(9, "Reds")))
mean_agg_cases <- ggplot()+
  geom_sf(data = eth_sf2, aes(fill = mean_cases), col = "black", lwd = .3) +
  geom_sf(data = m_0, col = "black", lwd = .8, alpha = 0) + 
  # scale_fill_gradientn(colours = mypalette(120), limits = c(0,120), breaks = seq(0,120, 20)) +
    scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Estimated aggregated cases, mean (x1000)") +
  theme_void(base_family = "Arial") +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5)) +  annotation_scale(location = "br", width_hint = 0.5)  +
  annotation_north_arrow(location = "tl", which_north = "true",
        pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
        style = north_arrow_fancy_orienteering)


LL_agg_cases <- ggplot()+
  geom_sf(data = eth_sf2, aes(fill = LL_cases), col = "black", lwd = .3) +
  geom_sf(data = m_0, col = "black", lwd = .8, alpha = 0) + 
  # scale_fill_gradientn(colours = mypalette(120), limits = c(0,120), breaks = seq(0,120, 20)) +
  scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Estimated aggregated cases, Lower limit (x1000)") +
  theme_void(base_family = "Arial") +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5)) +  annotation_scale(location = "br", width_hint = 0.5)

UL_agg_cases <- ggplot()+
  geom_sf(data = eth_sf2, aes(fill = UL_cases), col = "black", lwd = .3) +
  geom_sf(data = m_0, col = "black", lwd = .8, alpha = 0) + 
  scale_fill_gradientn(colours = mypalette(120), limits = c(0,120), breaks = seq(0,120, 20)) +
  # scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Estimated aggregated cases, Upper limit (x1000)") +
  theme_void(base_family = "Arial") +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5)) +  annotation_scale(location = "br", width_hint = 0.5)

library(patchwork)
p_agg <- mean_agg_cases + (UL_agg_cases/LL_agg_cases) + plot_annotation(tag_levels = "A")
p_agg
```


### Aggregated map on district level
```{r}
mean_prev <- exactextractr::exact_extract(r_prev_mean *100, eth_sf2, fun = "mean") # faster
eth_sf2$INLA_prev_mean <- mean_prev

sd_prev <- exactextractr::exact_extract(r_prev_sd*100, eth_sf2, fun = "mean") # faster
eth_sf2$INLA_prev_sd <- sd_prev

min_prev <- exactextractr::exact_extract(r_prev_mean *100, eth_sf2, fun = "min") # faster
eth_sf2$INLA_prev_min <- min_prev

max_prev <- exactextractr::exact_extract(r_prev_mean *100, eth_sf2, fun = "max")
eth_sf2$INLA_prev_max <- max_prev

eth_sf2$range <- eth_sf2$INLA_prev_max - eth_sf2$INLA_prev_min
```

```{r}
mypalette <- colorRampPalette((brewer.pal(9, "Reds")))
prev <- ggplot()+
  geom_sf(data = eth_sf2, aes(fill = INLA_prev_mean), col = "black", lwd = .3) +
  scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Mean prevalence (%)") +
  theme_void(base_family = "Arial") +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5))
```
+ Range
```{r}
range <- ggplot()+
  geom_sf(data = eth_sf2, aes(fill = range), col = "black", lwd = .3) +
  scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Range of Mean prevalence (%)") +
  theme_void(base_family = "Arial") +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5))
```

```{r, warning=FALSE, message=FALSE}
plot_grid(prev, range, labels = "AUTO")
```

### Relationship profile between environmental variables and the predicted prevalence
```{r}
data_p <- dp[, varlist]
data_p$pred_prev <- prev_mean * 100
# data_p$pred_prev_UL <- prev_ul
# data_p$pred_prev_LL <- prev_ll
# data_p$pred_prev_med <- prev_med

data_p_long <- data_p %>% pivot_longer(cols = names(data_p)[1:8],
                                       names_to = "covariates",
                                       values_to = "values") %>% 
  mutate(covariates = as_factor(covariates))    

levels(data_p_long$covariates) <- c("Slope", "Isothermality", "Precipitation seasonality",
                                    "Vegetation index", "Distance to the nearest river (km)", "Population density", "Flow accumulation", "Soil moisture (mm)")
```
```{r, warning=FALSE, message=FALSE}
p <- ggplot(data_p_long, aes(values, pred_prev)) +
  geom_smooth(method = "gam") + facet_wrap(~covariates, scales = "free", nrow = 2) + # default method is gam
  theme_bw(base_family = "Arial", base_size = 12) +
  xlab("Values") + ylab("Onchocerciasis prevalence")
p
```

### Plotting the spatial field
+ Thank to this [blog](https://ourcodingclub.github.io/tutorials/inla/)

```{r}
points.em <- mesh1$loc

stepsize <- 0.008 # This is given in coordinates unit (in this case this corresponds to 1km)
east.range <- diff(range(points.em[,1]))  # calculate the length of the longitude
north.range <- diff(range(points.em[,2])) # calculate the length of the latitude

nxy <- round(c(east.range, north.range)/stepsize)  # Calculate the number of cells in the x and y ranges

# Project the spatial field on the mesh vertices using the inla.mesh.projector() function
projgrid <- inla.mesh.projector(mesh1,
                                xlim = range(points.em[,1]),
                                ylim = range(points.em[,2]),
                                dims = nxy)
xmean <- inla.mesh.project(projgrid,
                           res$summary.random$spatial.field$mean)
xsd <- inla.mesh.project(projgrid,
                         res$summary.random$spatial.field$sd)

xmean2 <- t(xmean)
xmean3 <- xmean2[rev(1:length(xmean2[,1])),]
xmean_ras <- raster(xmean3,
                    xmn = range(projgrid$x)[1], xmx = range(projgrid$x)[2],
                    ymn = range(projgrid$y)[1], ymx = range(projgrid$y)[2],
                    crs = CRS("+proj=longlat +datum=WGS84 +no_defs"))

xsd2 <- t(xsd)
xsd3 <- xsd2[rev(1:length(xsd2[,1])),]
xsd_ras <- raster(xsd3,
                  xmn = range(projgrid$x)[1], xmx =range(projgrid$x)[2],
                  ymn = range(projgrid$y)[1], ymx =range(projgrid$y)[2],
                  crs = CRS("+proj=longlat +datum=WGS84 +no_defs"))
xmean_ras2 <- mask(xmean_ras, m)
xsd_ras2 <- mask(xsd_ras, m)
```

```{r, warning=FALSE, message=FALSE}
my.palette.var <- brewer.pal(n = 9, name = "BuPu")
my.palette.heat <- rev(brewer.pal(n = 9, name = "RdYlBu"))

p_mean <- tm_shape(xmean_ras2) + 
  tm_raster(title="Spatial field (Mean)", alpha = 1, palette = my.palette.heat, legend.is.portrait = T, style = "cont", midpoint = NA)+
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c(.1, .75)) +
  # tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c(.6, .1)) +
  tm_layout(legend.outside = FALSE, legend.position = c(.61, .65), frame = FALSE, fontfamily = "Arial")


p_var <- tm_shape(xsd_ras2) + 
  tm_raster(title="Spatial field (SD)", alpha = 1, palette = my.palette.var, legend.is.portrait = T, style = "cont", midpoint = NA)+
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  # tm_compass(type = "arrow", position = c(.1, .75)) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c(.6, .12)) +
  tm_layout(legend.outside = FALSE, legend.position = c(.61, .65), frame = FALSE, fontfamily = "Arial")

spatial_field <- tmap_arrange(p_mean, p_var, ncol = 2, widths = c(1,1), heights = 1)

spatial_field
```

### Plotting observed and the predicted values from the model
```{r}
data$PREV <- data$CASES/data$N  *100
pred_stack <- stack(r_prev_mean, r_prev_ll, r_prev_ul)
names(pred_stack) <- c("predicted_mean_prev", "pred_LL_prev", "pred_UL_prev")
pred_eth <- raster::extract(pred_stack, eth_prev2[c("LONG","LAT")], na.rm = TRUE, df = TRUE)


data2 <- as.data.frame(cbind(data, pred_eth)) 
data2$predicted_mean_prev <- data2$predicted_mean_prev *100
cor(data$PREV, pred_eth$predicted_mean_prev, method = "spearman")

data2 <- data2 %>% dplyr::select(PREV, predicted_mean_prev, pred_LL_prev,
                                                         pred_UL_prev) %>% mutate(abs_difference = (abs(PREV-predicted_mean_prev)))
```

```{r}
mypalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))

p <-ggplot(data2) +
  geom_point(mapping = aes(PREV , predicted_mean_prev, col = abs_difference)) +
  geom_abline(slope=1, intercept=0, linetype = "dashed") +
  ggpubr::stat_cor(mapping = aes(PREV, predicted_mean_prev), data = data2, method = "spearman", label.x = 50, label.y = 5) +
  scale_color_gradientn(colours = mypalette(100)) + 
  theme_bw(base_family = "Arial", base_size = 12) +
  xlab("Observed prevalence(%)") +
  ylab("Predicted prevalence (%)") + labs(color = "Absolute error") +
  theme(legend.position = "bottom", legend.key.size = unit(1, 'cm'),                                                                       #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm')) #+ facet_wrap(~variables)
p
```

## Running the regular binomial model
```{r, eval = FALSE}
res <- inla(predterms,
            family = "binomial", Ntrials = numtrials,
            data = inla.stack.data(datastack, spde = spde),
            control.family = list(link = "logit"),
            control.compute = list(dic = TRUE, waic = TRUE,
                                   cpo = TRUE, config = TRUE,
                                   openmp.strategy="huge"),
            control.predictor = list(
              compute = TRUE, link = 1,
              A = inla.stack.A(datastack)
            )
)
```

```{r}
res <- readRDS("data/inla_point_res_5km_mesh5_binomial.rds")
```

```{r}
# Mapping prediction ------------------------------------------------------
index <- inla.stack.index(stack = datastack, tag = "pred")$data

prev_mean <- res$summary.fitted.values[index, "mean"]
prev_ll <- res$summary.fitted.values[index, "0.025quant"]
prev_ul <- res$summary.fitted.values[index, "0.975quant"]
prev_sd <-  res$summary.fitted.values[index, "sd"]

summary(res$summary.fitted.values[index,])
load("data/220508_INLA_model_run_5km_binom.RData")
```

```{r, eval = FALSE}
r_prev_mean <- raster_prev(prev_mean = prev_mean)
r_prev_ul <- raster_prev(prev_ul)
r_prev_ll <- raster_prev(prev_ll)
r_prev_sd <- raster_prev(prev_sd)
```

+ Mean prevalence
```{r}
pal <- hcl.colors(100, palette = "Reds", rev =TRUE)

p_mean <- tm_shape(r_prev_mean*100) + 
  tm_raster(title="Binomial: Mean Prevalence (%)", alpha = 1, palette = pal, legend.is.portrait = F, style = "cont", midpoint = 50) +
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p_mean
```

+ Prevalence upper limit
```{r}
p_UL <- tm_shape(r_prev_ul*100) + 
  tm_raster(title="Binomial: Prevalence (%), UL", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 50) +
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p_UL
```

+ Prevalence lower limit
```{r}
p_LL <- tm_shape(r_prev_ll*100) + 
  tm_raster(title="Binomial: Prevalence (%), LL", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 50) +
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p_LL
```

+ Prevalence uncertainty
```{r}
pal <- hcl.colors(100, palette = "Blues", rev =TRUE)
p <- tm_shape(r_prev_sd*100) + 
  tm_raster(title="Prevalence(SD)", alpha = .8, palette = pal, legend.is.portrait = T, style = "cont") +
  tm_shape(m_0)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_shape(data_sf) + 
  tm_symbols(col = "black", shape = 3, alpha = .5, size = .07, legend.shape.show = T) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
p
```


```{r, eval = FALSE}
# Mapping exceedance probabilities ----------------------------------------

index <- inla.stack.index(stack = datastack, tag = "pred")$data
marg <- res$marginals.fitted.values[index][[1]]
1 - inla.pmarginal(q = 0.20, marginal = marg)

## P(Prevalence>20%)
excprob_20 <- sapply(res$marginals.fitted.values[index],
                  FUN = function(marg){1-inla.pmarginal(q = 0.20, marginal = marg)})
```

```{r}
pal <- hcl.colors(100, palette = "RdYlBu", rev =TRUE)

excprob_20_map <- tm_shape(raster_excprob_20) + 
  tm_raster(title="P (nodule prevalence > 20%)", alpha = 1, palette = pal, legend.is.portrait = T, style = "cont", midpoint = 0.5) +
  tm_shape(m)+
  tm_borders(lty = 1, lwd = 2, col = "black", alpha = .9) +
  tm_compass(type = "arrow", position = c("left", "top")) +
  tm_scale_bar(breaks = c(0, 100, 200, 400), text.size = .8, position = c("right", "bottom")) +
  tm_layout(legend.outside = FALSE, legend.position = c("right", "top"), frame = FALSE, fontfamily = "Arial")
excprob_20_map
```

